---
title: "NE03_Scrublet_raw"
author: "Spyros Darmanis"
date: "3/12/2019"
output: html_document
---

```{r Set directory}
rm(list=ls())
dir <- "/Users/spyros/Documents/Labwork/Projects/Biohub/NE_Lung/"
```

Load data and metadata
```{r Write data and metadata tables}
data_adult <- read.table(paste(dir,"input/secondary_input/GC_table_adult_NE_renamed.tsv",sep=""), check.names = F)
meta_adult_new <- read.table(paste(dir,"input/secondary_input/Metadata_table_adult_NE_renamed.tsv",sep=""), check.names = F)
```

Create experiment ID
Barcode XXXXNXXKZZ
X: ignore 
N: project ID
K: experiment ID
ZZ:plate ID (within each experiment)
```{r Separate experiment IDs}
meta_adult_new$experiment.ID <- substr(meta_adult_new$plate, 8,8)
table(meta_adult_new$experiment.ID)
exps <- unique(meta_adult_new$experiment.ID)
```

```{r Divide data by experiment and write matrix to Scrublet output}
require(Matrix)
# Get all experiment IDs
# exps <- c(6,7,8)
list.names <- list()
for(i in 1:length(exps)){
  # Subset cells and adata for each experiment ID
  cells <- row.names(meta_adult_new)[which(meta_adult_new$experiment.ID==exps[i])]
  list.names[[i]] <- cells
  # paste("Ncells for experiment ID", exps[i], "=", length(cells))
  message(paste("Ncells for experiment ID", exps[i], "=", length(cells)))
  data.temp <- as.matrix(data_adult[,cells])
  # Convert to sparse format 
  data.temp <- Matrix(data.temp, sparse=T)
  colnames(data.temp) <- cells
  #write.table(data.temp, file=paste("/Volumes/GoogleDrive/My Drive/scrublet/NE_analysis/input/","Exp.ID.", exps[i],".tsv" ,sep=""))
  writeMM(data.temp, file=paste(paste(dir,"input/scrublet/",sep=""),"Exp.ID.", exps[i],".mtx" ,sep=""))
}
names(list.names) <- exps
rm(data.temp)

```

Run script NE_scrublet_basics to generate doublet score for each cell in each run 
Import doublet scores for each cell 
```{r Import and combine doublet scores}
temp.6 <- read.table(paste(dir,"input/scrublet/NE_Lung_Exp.ID_6_doubletScore.csv", sep=""), sep=",")
colnames(temp.6) <- list.names$`6`
temp.7 <- read.table(paste(dir,"input/scrublet/NE_Lung_Exp.ID_7_doubletScore.csv", sep=""), sep=",")
colnames(temp.7) <- list.names$`7`
temp.8 <- read.table(paste(dir,"input/scrublet/NE_Lung_Exp.ID_8_doubletScore.csv", sep=""), sep=",")
colnames(temp.8) <- list.names$`8`
temp.all <- t(cbind(temp.6, temp.7,temp.8))
rm(list=c("temp.6", "temp.7","temp.8"))
# Add to Seurat metadata 
meta_adult_new$scrublet_score <-  temp.all[row.names(meta_adult_new),]
```

Import doublet scores for each cell 
```{r Import and combine doublet classification}
temp.6 <- read.table(paste(dir,"input/scrublet/NE_Lung_Exp.ID_6_doubletpred.csv", sep=""), sep=",")
colnames(temp.6) <- list.names$`6`
temp.7 <- read.table(paste(dir,"input/scrublet/NE_Lung_Exp.ID_7_doubletpred.csv", sep=""), sep=",")
colnames(temp.7) <- list.names$`7`
temp.8 <- read.table(paste(dir,"input/scrublet/NE_Lung_Exp.ID_8_doubletpred.csv", sep=""), sep=",")
colnames(temp.8) <- list.names$`8`
temp.all <- t(cbind(temp.6, temp.7,temp.8))
rm(list=c("temp.6", "temp.7","temp.8"))
# Add to Seurat metadata 
meta_adult_new$scrublet_doublet <-  temp.all[row.names(meta_adult_new),]
meta_adult_new$scrublet_doublet[meta_adult_new$scrublet_doublet=="True"] <- T
meta_adult_new$scrublet_doublet[meta_adult_new$scrublet_doublet=="False"] <- F
```
Do things make sense ? 

```{r Table}
table(meta_adult_new$scrublet_doublet)
```


Save data and metadata 
```{r }
write.table(data_adult, paste(dir,"input/secondary_input/GC_table_adult_NE_renamed_after_Scrublet.tsv",sep=""))
write.table(meta_adult_new, paste(dir,"input/secondary_input/Metadata_table_adult_NE_renamed_after_Scrublet.tsv",sep=""))
```

Remove doublets and save the tables again 
```{r }
cells <- row.names(meta_adult_new)[which(meta_adult_new$scrublet_doublet==F)]
data_adult_ND <- data_adult[,cells]
# dim(data_adult_ND)
# dim(data_adult)
meta_adult_new_ND <- meta_adult_new[cells,]
# dim(meta_adult_new_ND)
```

Save data and metadata 
```{r }
write.table(data_adult_ND, paste(dir,"input/secondary_input/GC_table_adult_NE_renamed_ND.tsv",sep=""))
write.table(meta_adult_new_ND, paste(dir,"input/secondary_input/Metadata_table_adult_NE_renamed_ND.tsv",sep=""))
```




