---
title: "NE01_Import_and_create_Seurat"
author: "Spyros Darmanis"
date: "3/12/2019"
output: html_document
---

Import master data tables
```{r Import}
#master.data <- read.table("/Users/spyros/Documents/Labwork/Projects/Biohub/NE_Lung/master_input/CIRM_master_expression_count_table.tsv", row.names = 1)
```


Subset cells of interest 
```{r Subset}
# data <- master.data[,grep("zsGreen", colnames(master.data))]
# dim(data)
# write.table(data, file="/Users/spyros/Documents/Labwork/Projects/Biohub/NE_Lung/input/GC_table_all_zsGreen.tsv")
```

Set directory
This is used to save within the repo
All paths can be changed accordingly
```{r Set directory}
dir <- "/Users/spyros/Documents/Labwork/Projects/Biohub/NE_Lung/"
```

```{r Import zsGreen data}
#rm(list=ls())
data <- read.table(paste(dir,"input/GC_table_all_zsGreen.tsv",sep=""), row.names = 1, header = T)
dim(data)
```

Generate metadata file for cells of interest
```{r Create metadata}
metadata <- as.data.frame(matrix(nrow=ncol(data), ncol=2))
row.names(metadata) <- colnames(data)
row.names(metadata) <- gsub("-zsGreen-1-","-zsGreen-1",row.names(metadata))
dim(metadata)
```

Change cell names for the mis-labelled library
```{r Correct demux}
cor_demux <- read.csv(paste(dir,"input/correct_demux.csv",sep=""), header=T)
incor_demux  <- read.csv(paste(dir,"input/incorrect_demux.csv",sep=""), header=T)
# Correct the names of the cells in the metadata file 
metadata[,"initial_name"] <- row.names(metadata)
metadata[,"actual_name"] <- row.names(metadata)
# Given name -> incorrect barcode -> correct name 
for(i in 1:nrow(metadata)){
  # get cell name. That name was assigned to a wrong cell. 
  cell <- metadata$initial_name[i]
  # which index corresponds to that cell in the wrong sample sheet ? 
  index <- as.character(incor_demux[which(incor_demux[,2]==cell),"Index"])
  if(length(index)!=0){
  # Which cell got the actual index instead of the correct cell ? 
  metadata[i,"actual_name"] <-  as.character(cor_demux[which(as.character(cor_demux$Index)==index),"SampleID"])}
}
```

Now we need to rename the data fields as well 
Initial name in metadata matches data column names 
Replace colnames with actual name 
```{r Fix names}
colnames(data) <- metadata$actual_name
row.names(metadata) <- metadata$actual_name
metadata <- metadata[,-c(3,4)]
colnames(metadata) <- c("plate", "well")
list <- strsplit(row.names(metadata), ".", fixed = T)
for(i in 1:length(list)){
  metadata$plate[i] <- list[[i]][2]
  metadata$well[i] <- list[[i]][1]
}
rm("list")
head(metadata)
```

Populate metadata fields

Clean up GC table
```{r clean data}
# Remove stat rows that contain the special character "_"
data_stats <- data[grep("no_feature", row.names(data)):nrow(data),]
data_genes <- data[-c(grep("no_feature", row.names(data)):nrow(data)),]
# Remove ERCC rows 
ercc_rows <- grep("ERCC-", row.names(data_genes))
data_ercc <- data_genes[ercc_rows,]
data_genes <- data_genes[-ercc_rows,]
data_genes[1:5,1:5]
```

```{r populate metadata, Nreads}
metadata[,"ERCC_reads"] <- colSums(data_ercc) 
metadata[,"Non_ERCC_reads"] <- colSums(data_genes) 
metadata[,"ERCC_to_non_ERCC"] <- metadata[,"ERCC_reads"]/metadata[,"Non_ERCC_reads"]
hist(metadata$Non_ERCC_reads, 200, main="N reads all cells")
```

```{r populate metadata, Ngenes}
metadata[,"Genes_detected"] <- NA
for(i in 1:ncol(data_genes)){
  a <- which(row.names(metadata)==colnames(data_genes)[i])
  metadata$Genes_detected[a] <- length(which(data_genes[,i] > 1))
}
hist(metadata$Genes_detected, 200)
```

```{r populate metadata, Age}
require(ggthemes)
metadata[,"Age"] <- NA 
metadata[grep("10005001", metadata$plate),"Age"] <- "E14"
metadata[grep("10005002", metadata$plate),"Age"] <- "E14"
metadata[grep("10005003", metadata$plate),"Age"] <- "E17.5"
metadata[grep("10005004", metadata$plate),"Age"] <- "E19.5"
metadata[grep("10005005", metadata$plate),"Age"] <- "E15"
metadata[grep("10005006", metadata$plate),"Age"] <- "PN21"
metadata[grep("1000500608", metadata$plate),"Age"] <- "PN120"
metadata[grep("1000500609", metadata$plate),"Age"] <- "PN120"
metadata[grep("10005007", metadata$plate),"Age"] <- "PN120"
metadata[grep("10005008", metadata$plate),"Age"] <- "PN90"
metadata$Age <- as.factor(metadata$Age)
# And color 
metadata[,"Age_color"] <- tableau_color_pal("tableau20")(20)[metadata$Age]
table(metadata$Age)
```

```{r Subset adult cells}
meta_adult <- metadata[grep("PN", metadata$Age),]
```

Add experiment ID from plate barcode and add color vector per plate 
```{r Plate and experiment IDs and color vectors}
meta_adult$experiment.ID <-  substr(meta_adult$plate,start = 8,stop = 8)
meta_adult[,"plate_color"] <- tableau_color_pal("tableau20")(20)[as.factor(meta_adult$plate)]
```


Change all cell names to well_plate
```{r Rename cells to well_plateID in data and metadata}
# Data
list <- do.call(rbind, strsplit(colnames(data_genes), ".", fixed=T))
colnames(data_genes) <- paste(list[,1], list[,2], sep="_")
# Metadata
list <- do.call(rbind, strsplit(row.names(meta_adult), ".", fixed=T))
row.names(meta_adult) <- paste(list[,1], list[,2], sep="_")
```

Subset adult data 
```{r Subset adult data}
data_adult <- data_genes[,row.names(meta_adult)]
dim(data_adult)
```

Add annotation of cell type and doublet from previous workflow 
```{r consolidate annotations}
meta.old <- read.table(paste(dir,"input/Metadata_adult_lung_OLD.txt",sep=""), header=T, row.names = 1)
# Rename cells of old metadata
list <- do.call(rbind, strsplit(row.names(meta.old), "-"))
row.names(meta.old) <- paste(list[,1], list[,2], sep="_")
# Join old and new metadata
meta_adult$cell.id <- row.names(meta_adult)
meta.old$cell.id <- row.names(meta.old)
# Join
dim(meta_adult)
dim(meta.old)
t <- merge(meta_adult, meta.old, all=T, by="cell.id")
dim(t)
```

```{r Clean combined metadata keeping columns we need}
meta_adult_new <- t[,c("cell.id", "plate.x", "well.x", "ERCC_reads.x", "Non_ERCC_reads.x", "ERCC_to_non_ERCC.x", 
  "Genes_detected.x", "Age.x", "Age_color.x", "experiment.ID", "Cluster_2d", "cell.type", "FSC", "SSC", "APC_Lineage",
  "DAPI","PE.Cy7_Epcam","zsGreen", "sum_class","pulse.width")]
table(meta_adult_new$cell.type)
table(is.na(meta_adult_new$cell.type))
```

```{r Change names}
row.names(meta_adult_new) <- meta_adult_new$cell.id
```

Write data and metadata
```{r Write data and metadata tables}
write.table(data_adult, paste(dir,"input/GC_table_adult_NE.tsv",sep=""))
write.table(meta_adult_new, paste(dir,"input/Metadata_table_adult_NE.tsv",sep=""))
```


Create the Seurat object with all the data and save Seurat object
```{r create Seurat}
require(Seurat)
tiss <- CreateSeuratObject(raw.data = data_adult, project = "NE_Lung", meta.data = meta_adult_new, 
                           display.progress = T, min.cells = 0,min.genes = 0)
tiss <- AddMetaData(object = tiss, metadata = meta_adult_new, col.name = "Non_ERCC_reads.x")
save(tiss, file = paste(dir,"input/NE_Lung_Seurat.RData",sep=""))
```
